---
title: "Simulate Mittag-Leffler"
author: "Peter Straka"
date: "18/10/2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Generating Stables

We first generate random variables $X$ with the Laplace transform

$$\mathcal L[X](\lambda) = \mathbf E e^{-\lambda X} = e^{-\lambda^\alpha}$$ 

where $\alpha \in (0,1)$. 
The package `stabledist` with option `pm=1` runs with the parametrisation
as in Samorodnitsky & Taqqu: 
\begin{align*}
\mathcal F[X](\theta) = \mathbf E e^{i\theta X} = \exp \left\lbrace -\sigma^\alpha |\theta|^\alpha \left(1 - i\beta({\rm sign} \theta) \tan \frac{\alpha \pi}{2}\right) + i \mu \theta \right\rbrace.
\end{align*}

Hence we aim to find $\alpha, \beta, \sigma, \delta$ such that 
$$\mathcal L[X](-i\theta) = \mathcal F[X](\theta), \quad \theta > 0.$$

Now see that 
\begin{multline*}
e^{-(-i\theta)^\alpha}
= \exp\left(-(\theta e^{-i \pi / 2})^\alpha\right)
= \exp\left(-\left(\theta^\alpha e^{-i\alpha \pi/2}\right)\right)
= \exp\left(-\theta^\alpha\left(\cos \frac{\alpha\pi}{2} - i \sin\frac{\alpha\pi}{2}\right)\right)
\\
= \exp\left(-\theta^\alpha \cos\frac{\alpha\pi}{2}\left(1 - i \tan \frac{\alpha\pi}{2}\right)\right)
=  \mathcal F[X](\theta), \quad \theta > 0
\end{multline*}
whenever $\sigma^\alpha = \cos\frac{\alpha\pi}{2}, \beta = 1, \mu = 0$. 

```{r}
require(stabledist)
n = 500
# set parameters: 
alpha = 0.4
beta = 1
sigma = (cos(alpha * pi / 2))^(1/alpha)
mu = 0
# generate stables: 
y <- rstable(n = n, alpha = alpha, beta = beta, gamma = sigma, delta = mu, pm = 1)
# exponentials:
x <- rexp(n = n, rate = 1)
# Mittag-Lefflers:
w <- y * x^(1/alpha)
```

```{r}
source("../MittagLefflerEstimation.R", local=TRUE)
ml.par.est.delta(10000*w,0.95)
```

